Grove: A Convergent Collaborative Structure-Editor Calculus

### 2020-07-03 grove-paper.tex
Updates from Overleaf

%% Packages
\usepackage{booktabs}
\usepackage{subcaption}
\usepackage{tikz}

@@

\usetikzlibrary{external}
%\tikzexternalize % activate!
%\usepackage{lua-visual-debug}

@@

@@

\subsection{User Interface}

TODO: include mocked-up screenshots

@@

\begin{figure}
\centering
TODO: commutativity diagram of screenshots with two cursors
\caption{TODO:Caption}
\label{fig:my_label}
\end{figure}

\begin{figure}
\centering
\begin{tikzpicture}
% TODO: texttt
% TODO: position the different graphs
% TODO: ballanced trees
% TODO: use tikz graph notation? tree notation?
\path (0cm,0cm) graph[tree layout] {
  + -> {
    3[as={\texttt{3}}], % TODO: put a cursor (green) here
    4 % TODO: put a cursor (red) here
  }
};
\path (-3cm,-2cm) graph[tree layout] {
  + -> {
    13, % TODO: put a cursor here
    4
  }
};
\path (3cm,-2cm) graph[tree layout] {
  + -> {
    3,
    14 % TODO: put a cursor here
  }
};
\path (0cm,-4cm) graph[tree layout] {
  + -> {
    13, % TODO: put a cursor here
    14 % TODO: put a cursor here
  }
};
\end{tikzpicture}
\caption{Scenario A: Editing Different Parts of the Code}
\label{fig:Scenario-A}
\end{figure}

@@

\begin{figure}
\centering
\begin{tikzpicture}
% TODO: texttt
% TODO: position the different graphs
% TODO: ballanced trees
% TODO: use tikz graph notation? tree notation?
\path (0cm,0cm) graph[tree layout] {
  lam[as={$\mathrm{\lambda}_{\texttt{x}}$}] -> {
    * -> {
      2,
      x
    }
  }
};
\path (-3cm,-2cm) graph[tree layout] {
  lam[as={$\mathrm{\lambda}_{\texttt{x}}$}] -> {
    + -> {
      1,
      * -> {
        2,
        x
      }
    }
  }
};
\path (3cm,-2cm) graph[tree layout] {
  lam[as={$\mathrm{\lambda}_{\texttt{x}}$}] -> {
    * -> {
      3,
      x
    }
  }
};
\path (0cm,-4cm) graph[tree layout] {
  lam[as={$\mathrm{\lambda}_{\texttt{x}}$}] -> {
    + -> {
      1,
      * -> {
        3,
        x
      }
    }
  }
};
\end{tikzpicture}
\caption{Scenario B: Edits to Nested Parts of the Code}
\label{fig:Schenario-B}
\end{figure}

@@

@@

\begin{figure}
\centering
\begin{tikzpicture}
% TODO: texttt
% TODO: position the different graphs
% TODO: ballanced trees
% TODO: use tikz graph notation? tree notation?
\path (0cm,0cm) graph[tree layout] {
  lam[as={$\mathrm{\lambda}_{\texttt{x}}$}] -> {
    x
  }
};
\path (-3cm,-2cm) graph[tree layout] {
  lam[as={$\mathrm{\lambda}_{\texttt{x}}$}] -> {
    + -> {
      1,
      x
    }
  }
};
\path (3cm,-2cm) graph[tree layout] {
  lam[as={$\mathrm{\lambda}_{\texttt{x}}$}] -> {
    * -> {
      2,
      x
    }
  }
};
\path (2cm,-4cm) graph[tree layout] {
  lam[as={$\mathrm{\lambda}_{\texttt{x}}$}] -> {
    + -> {
      1,
      * -> {
        3,
        x
      }
    }
  }
};
\path (-2cm,-4cm) graph[tree layout] {
  lam[as={$\mathrm{\lambda}_{\texttt{x}}$}] -> {
    * -> {
      3,
      + -> {
        1,
        x
      }
    }
  }
};
\end{tikzpicture}
\caption{Scenario C: Direct Conflicts}
\label{fig:Scenario-C}
\end{figure}

@@

@@

@@

@@

\subsection{Scenario E: Restoring Deleted Code}
If only one restore, then it does what you expect (basically a cut and paste)

@@

@@

Compared to the core language, our implementation lacks support for lists.

>>>>

%% Packages
\usepackage{booktabs}
\usepackage{subcaption}
\usepackage[pdf]{graphviz}

@@

\usetikzlibrary{external} % Needs to be separately enabled
%\tikzexternalize % Enable externalization
%\usepackage{lua-visual-debug}

@@

TODO: gray-scale unchanged parts of diagrams

@@

TODO: annotate vertexes and edges with ids (gray superscript?)

Single User Intuition:
 - manipulating trees,
 - holes

Multi User Intuition:
 - vertexes have unique ids
 - edits create or delete edges in the tree,
 - vertexes are automatically created when an edge goes to it

Deeper intuition:
 - edges have unique ids
 - once deleted an edge can never be recreated (though, a new edge (with a new id) can be created in the same spot as the original)

\subsection{Scenario X: Holes in the Tree}
\label{sub:Holes in the Tree}

\begin{figure}
\centering
\digraph[]{simple-tree-with-hole}{
  node [shape=Mrecord,fontsize=11,ranksep=0];
  edge [arrowhead=vee,fontsize=11,weight=2];
  rankdir = LR;
  {rank=min; n0 [shape=point]};
  n0 [label="{0: Root_root|{<root> root}}",style=filled,fillcolor=black];
  n3 [label="{3: Exp_plus|{<left> left|<right> right}}",style=filled,fillcolor=white];
  n6 [label="{6: Exp_num(2)|{}}",style=filled,fillcolor="#ddffdd"];
  n3:left -> n6 [color=black,label="7",edgeURL="#",edgetooltip="id: 7\nsource: 3\nindex: Lang.Index.Exp_plus_left\ntarget: 6",labeltooltip="id: 7\nsource: 3\nindex: Lang.Index.Exp_plus_left\ntarget: 6"];
  n0:root -> n3 [color=black,label="4",edgeURL="#",edgetooltip="id: 4\nsource: 0\nindex: Lang.Index.Root_root_root\ntarget: 3",labeltooltip="id: 4\nsource: 0\nindex: Lang.Index.Root_root_root\ntarget: 3"]}
\begin{tikzpicture}
% TODO: texttt
% TODO: position the different graphs
\path (0cm,0cm) graph[tree layout] {
  + -> {
    2[as={\texttt{2}}],
    % <empty>
  }
};
\end{tikzpicture}
\caption{Scenario X: Holes in the Tree}
\label{fig:Scenario X: Holes in the Tree}
\end{figure}

DIAGRAM: 2 + HOLE

NEEDS UI EXAMPLE

@@

\begin{figure}
\centering
TODO: commutativity diagram of screenshots with two cursors
\caption{TODO:Caption}
\label{fig:Screenshot0}
\end{figure}

\begin{figure}
\centering
\begin{tikzpicture}
% TODO: texttt
% TODO: position the different graphs
\path (0cm,0cm) graph[tree layout] {
  + -> {
    3[as={\texttt{3}}], % TODO: put a cursor (green) here
    4 % TODO: put a cursor (red) here
  }
};
\path (-3cm,-2cm) graph[tree layout] {
  + -> {
    13, % TODO: put a cursor here
    4
  }
};
\path (3cm,-2cm) graph[tree layout] {
  + -> {
    3,
    14 % TODO: put a cursor here
  }
};
\path (0cm,-4cm) graph[tree layout] {
  + -> {
    13, % TODO: put a cursor here
    14 % TODO: put a cursor here
  }
};
\end{tikzpicture}
\caption{Scenario A: Editing Different Parts of the Code}
\label{fig:Scenario A: Editing Different Parts of the Code}
\end{figure}

@@

\begin{figure}
\centering
\begin{tikzpicture}
% TODO: texttt
% TODO: position the different graphs
\path (0cm,0cm) graph[tree layout] {
  lam[as={$\lambda_{\texttt{x}}$}] -> {
    * -> {
      2,
      x
    }
  }
};
\path (-3cm,-2cm) graph[tree layout] {
  lam[as={$\lambda_{\texttt{x}}$}] -> {
    + -> {
      1,
      * -> {
        2,
        x
      }
    }
  }
};
\path (3cm,-2cm) graph[tree layout] {
  lam[as={$\lambda_{\texttt{x}}$}] -> {
    * -> {
      3,
      x
    }
  }
};
\path (0cm,-4cm) graph[tree layout] {
  lam[as={$\lambda_{\texttt{x}}$}] -> {
    + -> {
      1,
      * -> {
        3,
        x
      }
    }
  }
};
\end{tikzpicture}
\caption{Scenario B: Edits to Nested Parts of the Code}
\label{fig:Schenario-B}
\end{figure}

TODO: REDO to MOVE a SUBTREE

NEEDS UI MOCKUP

@@

Contrast with a traditional diff

@@

\begin{figure}
\centering
\begin{tikzpicture}
% TODO: texttt
% TODO: position the different graphs
\path (0cm,0cm) graph[tree layout] {
  lam[as={$\lambda_{\texttt{x}}$}] -> {
    x
  }
};
\path (-3cm,-2cm) graph[tree layout] {
  lam[as={$\lambda_{\texttt{x}}$}] -> {
    + -> {
      1,
      x
    }
  }
};
\path (3cm,-2cm) graph[tree layout] {
  lam[as={$\lambda_{\texttt{x}}$}] -> {
    * -> {
      x,
      2
    }
  }
};
\path (0cm,-4cm) graph[tree layout] {
  lam[as={$\lambda_{\texttt{x}}$}] -> {
    + -> {
      1,
      x
    },
    * -> {
      x,
      2
    }
  }
};
\path (2cm,-8cm) graph[tree layout] {
  lam[as={$\lambda_{\texttt{x}}$}] -> {
    + -> {
      1,
      * -> {
        x,
        2
      }
    }
  }
};
\path (-2cm,-8cm) graph[tree layout] {
  lam[as={$\lambda_{\texttt{x}}$}] -> {
    * -> {
      + -> {
        1,
        x
      },
      2
    }
  }
};
\end{tikzpicture}
\caption{Scenario C: Direct Conflicts}
\label{fig:Scenario-C}
\end{figure}

@@

TODO: SHOW intermediate Steps (maybe no graph)
  Show user's perspective
  Accept one side = delete others (select this)
  Delete x in *
  Move + to x

UI Level Move = Delete and Add

@@

NEEDS UI MOCKUP

@@

\begin{figure}
\centering
\begin{tikzpicture}
% TODO: texttt
% TODO: position the different graphs
\path (0cm,0cm) graph[tree layout] {
  lam[as={$\lambda_{\texttt{x}}$}] -> {
    + -> {
      2,
      x
    }
  }
};
\path (-3cm,-2cm) graph[tree layout] {
  lam[as={$\lambda_{\texttt{x}}$}] -> {
    + -> {
      3,
      x
    }
  }
};
\path (3cm,-2cm) graph[tree layout] {
  lam[as={$\lambda_{\texttt{x}}$}] -> {
    hole
    },
  + -> {
    2,
    x
  }
};
\path (0cm,-4cm) graph[tree layout] {
  lam[as={$\lambda_{\texttt{x}}$}] -> {
    hole
    },
  + -> {
    3,
    x
  }
};
\end{tikzpicture}
\caption{Scenario D: Deletion of Code Being Edited}
\label{fig:Scenario-D}
\end{figure}

TODO: move into Scenario B

@@

\subsection{Scenario E: Restoring Deleted Code}

Combine with Scenario B (moving code):
  delete
  restore
  cut/paste (i.e., move) = delete then restore
  copy structure vs copy id
  paste structure vs paste id
    (show two or three options at paste time)

DIAGRAM: 1 * 2 + 3 + HOLE

DIAGRAM: HOLE + 3 + HOLE (1 * 2)

DIAGRAM: HOLE + 3 + 1 * 2

NEEDS UI MOCKUP

If only one restore, then it does what you expect (basically a cut and paste)

@@

\subsection{Scenario F: Cycles}

Synthetic cycles:

DIAGRAM: restoring into yourself (paste before delete?)

Organic cycles:

DIAGRAM:

NEEDS UI MOCKUP

@@

Compared to the core language, our implementation lacks support for lists and case expressions.

Optimizations: least fixed point
